---
title: "My first R Markdown document"
author: "Henrik"
output: word_document
---

## Getting started

The first thing to do in our analysis is load the R
packages that we'll use to do the work:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(lsr)
```

I am learning Markdown and I love it!

```{r hobbies}
hobbies <- c("diving", "learning", "swimming")
print(hobbies)
````

The next step is to load the data set itself. This time, we'll assume the data is stored in a csv (comma-delimited) format. We'll generally tend to use this from now on, because it's a format that nearly any language can read. It's usually a good idea to store data in general formats because that helps with sharing, so let's get into the habit of that.

```{r loaddatachunk}
loc <- here("mydata/gladlysurvey3.csv")
gdata3 <- read_csv(file = loc)
```
````{r piss}
view(gdata3, "name")
````

## Creating a histogram

Let's visualise our data. We'll look at much more awesome ways to visualise data later, but for now, let's just make a simple histogram of the years each person was born.

```{r gdhistogram}
hist(x=gdata$age)
```

Now make it nicer.

```{r gdhistogrammod}
hist(x=gdata$age,main="Ages of people in Gladly's Survey",xlab="Years Old",ylab="Number of people",col="blue",breaks=11,xlim=c(0,12))
`````````
Now we want to view the dataset. 

```{r showdatachunk}
summary(gdata)
```




