---
title: "Week 3 Day 1 analysis"
author: "Andy"
output: word_document
---

We'll begin by loading up the libraries we need, as always.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# loading the libraries
library(tidyverse)
library(here)
```

## Tables

First let's try making a table of people's colours.

```{r sampletablechunk}
# first the vector of colours
colours <- c("grey","brown","navy","white","white",
             "pink","pink","brown","yellow")
# now make a table out of it
colourtable <- table(colours)
# print the table
colourtable
```


Next let's try some cross-tabulation of colours by gender.

```{r samplectchunk, results="asis", echo=FALSE}
genders <- c("female","male","female","male","female","nb",
             "female","female","male")
gctable <- table(genders,colours)
gctable
cgtable <- table(colours,genders)
cgtable
```



```{r loadgladlydata, include=FALSE}
# loading the data
loc <- here("gladlysurvey3.csv")
gdata <- read_csv(file=loc)
```


Finally, let's make a table of some of Gladly's data. First, how many do we have of each species?

```{r specieschunk}
table(gdata$species)

```

## Pipes

Our first example.

```{r pipesexample}
# non-piped version
x <- c(1,5,3,2,5)
myMean <- mean(x)
roundedMean <- round(myMean)
absoluteRoundedMean <- abs(roundedMean)
absoluteRoundedMean

# piped version
absoluteRoundedMeanPipe <- c(1,5,3,2,5) %>% 
  mean() %>% 
  round() %>% 
  abs()
absoluteRoundedMeanPipe
```

Our second example.

```{r pipesexample2}
# draw 100 samples from a uniform distribution
x <- runif(n=100)

# non-piped version
sumX <- sum(x)
sqrtSumX <- sqrt(sumX)
rndSSX <- round(sqrtSumX,digits=2)
rndSSX

# piped version
rndSSXpipe <- x %>%
   sum() %>%
   sqrt() %>%
   round(digits=2)
rndSSXpipe
```

## Grouping

Let's use pipes and some of our dplyr functions to do some data manipulation. We're going to calculate the mean age for each gender using the `group_by()` and `summarise()` functions.

```{r genderagesummary}
gdata %>%
  group_by(gender) %>%
  summarise(mnAge = mean(age)) %>%
  ungroup()
```

### A simple example

This simple example dataset was made because it's so small we can show it on a slide, exactly as it happens.

```{r simpleExample}
# loading simple dataset.
# note we combined the here() and read_csv() into one line
# this is fine, either way works!
ds <- read_csv(file=here("simpleexample.csv"))

# create new summarised dataset called sumHappy
sumHappy <- ds %>%
    group_by(gender) %>%
    summarise(happiness = sum(happy,na.rm=TRUE)) %>%
    ungroup()
sumHappy
```


Now let's also calculate the standard deviation and the number of participants in each group. 

```{r genderageallsummary}
gdata %>%
  group_by(gender) %>%
  summarise(mnAge = mean(age),
            sdAge = sd(age),
            nAge = n()) %>%
  ungroup()
```


### Nesting

We can nest multiple variables.

```{r genderageallnestsummary}
gdata %>%
  group_by(gender,species) %>%
  summarise(mnAge = mean(age),
            sdAge = sd(age),
            nAge = n()) %>%
  ungroup()
```
A second example.

```{r example2, warning=FALSE, message=FALSE}
nestResult <- ds %>%
    group_by(gender,time) %>%
    summarise(meanJoy = mean(joy),
              n = n()) %>%
    ungroup()
nestResult
```


Now let's create some variables and do multiple operations on them.

```{r genderagemvsummary, warning=FALSE, message=FALSE}
genderSpeciesData <- gdata %>%
  group_by(gender,species) %>%
  summarise(mnAge = mean(age),
            sdAge = sd(age),
            nAge = n()) %>%
  ungroup()

genderSpeciesData %>%
  group_by(gender,species) %>%
  summarise(stdErr = sdAge/sqrt(nAge)) %>%
  ungroup()
```
