---
title: "Week 3 Exercises"
author: "Andrew Perfors"
output: word_document
---

```{r setup, include=FALSE}
# We'll begin by loading up the libraries and data we need, as always.
knitr::opts_chunk$set(echo = TRUE)

# loading the libraries
library(tidyverse)
library(here)

# loading the data
loc <- here("gladlysurvey3.csv")
gdata <- read_csv(file=loc)
```

## Day 1

### Tables

1. Create a summary table of the foods in Gladly’s data set and put it in a a variable called `foodtable`. Once you got it to work, knit this document to Word.

```{r day1tables1, include}
view(gdata)

foodSum <- c("food")

foodtable <- table(foodSum)



```

```{r sampletablechunck, include = TRUE}
# Jeg er litt trøtt

colours <- c("grey", "brown","navy","white","white","pink","pink","brown","yellow")
# now make a table out of it

colourtable <- table (colours)

# print the table
colourtable

# ny table
# ny column <- verdier

genders <- c("female","male","female","male","female","nb","female","female","male")

# kjønn table
gendertable <- table(genders)

#yes
gendertable

#crosstabulation
table(colours,genders)

```


2. Make a cross tab of age by ageCat. What do you think ageCat reflects? Do you think it is correctly calculated? [The latter two questions are thought questions asking you to think about and look at the data, no code necessary!]

```{r loadgladlydata, include=FALSE}
# loading the data
loc <- here ("gladlysurvey3.csv")
gdata <- read_csv(file=loc)

# species
table (gdata$species)

# data alder
table(gdata$age)

# ageCat

table("age", "cat")
```

3. Create a summary table of foods by species in the dataset. It will be kind of big. 

```{r specieschunck, echo=FALSE}

# species
table (gdata$species)

# data alder
table(gdata$age)

# ageCat

table("age", "cat")




```

### Pipes

1. The following code generates 100 random numbers from a normal distribution and calculates the mean of those numbers. Rewrite it using pipes and assign the output to a variable called meanNumsPipe.  

```{r day1pipes1}
# old code
nums <- rnorm(n=100)
meanNums <- mean(nums)

# new code put here
N <- 100

```

2. The following code generates 50 random numbers from a normal distribution with mean 100 and standard deviation 15. It then calculates the trimmed mean of those numbers. Rewrite it using pipes. Assign it to a new variable at the end. (A trimmed mean is the mean calculated on the dataset with a certain proportion of the ends removed. In this example, 20% is trimmed off each end. This is often more robust to extreme outliers than just taking the mean is.)

```{r day1pipes2}
# old code
nums <- rnorm(n=50,mean=100,sd=15)
trMeanNums <- mean(nums,trim=0.2)
trMeanNums

# new code
N <- 50

```

### Grouping

1. Calculate the mean age of each species in Gladly’s dataset and assign the answer to a variable called ageTibble. Make sure ageTibble is visible when you knit your document.

```{r day1grouping1, warning=FALSE, message=FALSE}

```

2. Group Gladly's data by species, and for each species calculate two variables: `mnCarrot` (which gives the mean answer to the carrot question) and `sumCarrot` (which gives the sum of everyone of that species answers to the carrot question). Note: you’ll have to use `na.rm` because there are some missing datapoints. Which of these do you think is more informative about how each species answers the question, and why?

```{r day1grouping2, warning=FALSE, message=FALSE}

```

3. What is the difference, if any, between `group_by(species,gender)` and `group_by(gender,species)`?

```{r day1grouping3, warning=FALSE, message=FALSE}

```


4. Group Gladly’s data by gender, and for each gender calculate `lowQn` (the 25th quantile), `hiQn` (the 75th quantile), and `median` (the 50th quantile) for age.

```{r day1grouping4, warning=FALSE, message=FALSE}

```

5. Challenge: in the last slide we calculated `stdErr` using the summary statistics calculated in a previous `group_by()` command. It’s possible to do this all in one set of pipes (i.e., without having to save the intermediate tibble). See if you can figure out how to do this. NOTE: if you can't do this, don’t worry! I just want you to practice playing around for yourself, but it's not necessarily intuitively obvious. 

```{r day1grouping5, warning=FALSE, message=FALSE}

```

