---
title: "RMHI/ARMP Problem Set 2"
author: 'Henrik Formoe 1575157 [Word Count: XX]'
output: word_document
---

Please put your answers here, following the instructions in the assignment description. Put your answers and word count tallies in the locations indicated; if none is indicated that means there is no word count for that question. Remember to knit as you go, and submit the knitted version of this on Canvas.

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
# We'll begin by loading up the libraries and data we need, as always.
knitr::opts_chunk$set(echo = TRUE)

# this deletes any variables that are in your environment
# useful so you don't get conflicts you're unaware of
rm(list=ls())

# loading the libraries
library(tidyverse)
library(here)
library(ggplot2)
library(lsr)
library(RColorBrewer)
library(car)

# loading datasets
ds <- read_csv(file=here("shiftsmonitoring.csv"))
dl <- read_csv(file=here("testdata.csv"))
dc <- read_csv(file=here("crimestats.csv"))
df <- read_csv(file=here("foodprices.csv"))
dh <- read_csv(file=here("healthratings.csv"))

# makes person variable in `ds` a factor
ds$person <- as.factor(ds$person)

# reorders so tables are in intuitive order
dc$crime <- factor(dc$crime,
                   levels=c("food","tech","vehicles","other"))
dc$year <- factor(dc$year,levels=c("previous","current"))
dh$exCat <- factor(dh$exCat,levels=c("low","high"))
dh$inCat <- factor(dh$inCat,levels=c("poor","rich"))
```

## Q1

**Q1a**

```{r q1a, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
# any code goes here
```


**Q1b**

```{r q1b, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*Note that the order of these assumptions (i.e., what assumption you put in 1 vs 2) does not matter!* 

*ASSUMPTION 1: Insert answer here [Word Count: N]*

*ASSUMPTION 2: Insert answer here [Word Count: N]*


**Q1c**

```{r q1c, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


**Q1d**

```{r q1d, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word count: N]*


## Q2 

**Q2a** 

```{r q2a, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


**Q2b** 

```{r q2b, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: (i) Insert answer here; (ii) Insert answer here; (iii) Insert answer here [Word Count: N]*


**Q2c**

```{r q2c, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: The standardised coefficient is ____.*


**Q2d** 

```{r q2d, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# this code is given
dl <- dl %>%
  mutate(lfbZnorm = (lfb - mean(lfb))/sd(lfb),
         foxyZnorm = (foxy - mean(foxy))/sd(foxy))

# any code goes here

```

*ANSWER: Insert answer here [Word Count: N]*


**Q2e**

```{r q2e, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


## Q3

**Q3a**

```{r q3a, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# this code is given
dcPlotTable <- dc %>%
  group_by(year,crime) %>%
  summarise(n = n()) %>%
  ungroup()

# any code goes here

```


**Q3b**

```{r q3b, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


**Q3c** 

```{r q3c, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


**Q3d**

```{r q3d, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


## Q4

```{r q4fig, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# plot the figure 
dfLong <- df %>%
  pivot_longer(cols=c(previous,current),names_to="year",
               values_to="price")
dfLong$year <- factor(dfLong$year,
                      levels=c("previous","current"))

dfSum <- dfLong %>%
  group_by(year) %>%
  summarise(mean=mean(price),
            sd=sd(price),
            n=n(),
            sdErr=sd/sqrt(n)) %>%
  ungroup()

dfLong %>%
  ggplot(mapping=aes(y=price,x=year,colour=year)) +
  geom_path(aes(group=names),colour="grey",
            linewidth=0.3) +
  geom_jitter(alpha=0.4,size=3,width=0.02,height=0) +
  geom_errorbar(data=dfSum,width=0.15,
                colour="black",
                mapping=aes(y=mean,
                            ymin = mean-sdErr,
                            ymax = mean+sdErr)) +
  theme_bw() +
  theme(legend.position="none") +
  scale_fill_brewer(palette="Set1") +
  scale_colour_brewer(palette="Set1") +
  labs(title="Price change over time",
       subtitle="Each dot is one kind of food item",
       y="Average price", x="Year")
```

**Q4a**

```{r q4a, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


**Q4b**

```{r q4b, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


## Q5

**Q5a**

```{r q5a, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```


**Q5b**

```{r q5b, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


## Q6

**Q6a**

```{r q6a, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: Insert answer here [Word Count: N]*


**Q6b**

```{r q6b, warning=FALSE, message=FALSE, fig.width=5, fig.height=5}
# any code goes here
```

*ANSWER: (i) Insert answer here; (ii) Insert answer here; (iii) Insert answer here [Word Count: N]*


**Q6c**

*ANSWER: Insert answer here [Word Count: N]*


## Q7

**Q7a**

*ANSWER: (i) ___ (ii) ___ (iii) ___ *


**Q7b**

*ANSWER: (i) Insert answer here; (ii) Insert answer here; (iii) Insert answer here [Word Count: N]*


## Q8

*ANSWER: Insert answer here.* 

