---
title: "RMHI/ARMP Problem Set 1"
author: 'Your name and ID goes here [Word Count: XX]'
---

Please put your answers here, following the instructions in the assignment description. Put your answers and word count tallies in the locations indicated; if none is indicated that means there is no word count for that question. Remember to knit as you go, and submit the knitted version of this on Canvas.

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
# We'll begin by loading up the libraries and data we need, as always.
knitr::opts_chunk$set(echo = TRUE)

# this deletes any variables that are in your environment
# useful so you don't get conflicts you're unaware of
rm(list=ls())

# loading the libraries
library(tidyverse)
library(here)
library(ggplot2)
library(RColorBrewer)

d <- read_csv(file=here("talentshow.csv"))
dd <- read_csv(file=here("talentshow_duration.csv"))

# these tibbles may be useful in some of the questions
d3b <- read_csv(file=here("d3b.csv"))
d6 <- read_csv(file=here("d6.csv"))
df <- read_csv(file=here("dfullMine.csv"))

# makes into factors
d$name <- as.factor(d$name)
dd$name <- as.factor(dd$name)
d$kind <- as.factor(d$kind)

# reorders so fun comes before competitive, makes figure nicer
d$level <- factor(d$level, levels=c("fun","competitive"))
dd$level <- factor(dd$level, levels=c("fun","competitive"))
```

## Q1 

**Q1a**

```{r q1a, warning=FALSE, message=FALSE}
# Put your code here

```

**Q1b**

```{r q1b, warning=FALSE, message=FALSE}
# Put your code here

```

*ANSWER: The most common talent was ______, with ____ performances.*


**Q1c** 

```{r q1c, warning=FALSE, message=FALSE}
# Put your code here

```


## Q2 

**Q2a**

```{r q2a, warning=FALSE, message=FALSE}
# Put your code here

```

**Q2b**

```{r q2b, warning=FALSE, message=FALSE}
# Put your code here

```

**Q2c**

*ANSWER: Put your answer here. [Word count: XX]* 


**Q2d**

```{r q2d, warning=FALSE, message=FALSE}
# Put your code here

```


## Q3

**Q3a**

```{r q3a, warning=FALSE, message=FALSE}
# Put your code here

```

**Q3b**

```{r q3b, warning=FALSE, message=FALSE}
# Put your code here

```

**Q3c**

```{r q3c, warning=FALSE, message=FALSE}
# optional code here

```

*ANSWER: Put your answer here. [Word count: XX]* 


**Q3d**

```{r q3d, warning=FALSE, message=FALSE}
# Put your code here

```

*ANSWER: The names of the individuals who broke Rule 1 (i.e., that everybody needs to participate in both fun and competitive levels) are ________. The names of the individuals who broke Rule 2 (i.e., that everybody must to do different kinds of talent at the fun and competitive levels) are ________.*


## Q4

**Q4a**

```{r q4a, warning=FALSE, message=FALSE}
# Put your code here

```

**Q4b**

```{r q4b, warning=FALSE, message=FALSE}
# Put your code here

```

**Q4c**

```{r q4c, warning=FALSE, message=FALSE}
# This code has been given to you, you just need to run it
dc <- cbind(d,dd)
head(dc)
```

*ANSWER: Put your answer here. [Word count: XX]* 


## Q5

**Q5a**

```{r q5a, warning=FALSE, message=FALSE}
# Put your code here

```

**Q5b** 

```{r q5b, warning=FALSE, message=FALSE}
# Put your code here

```

**Q5c**

*ANSWER: Put your answer here. [Word count: XX]* 


## Q6

**Q6a**

```{r q6a, fig.width=7, fig.height=4.5, warning=FALSE, message=FALSE}
# Put your code here

```

**Q6b**

*ANSWER: Put your answer here. [Word count: XX]* 


## Q7

**Q7a**

```{r q7a, fig.width=8, fig.height=6, warning=FALSE, message=FALSE}
# Put your code here

```

**Q7b**

*ANSWER: (1) Describe one new thing here. (2) Describe other new thing here. [Word count: XX]*


**Q7c**

*ANSWER: Put your answer here. [Word count: XX]* 


## Q8

*ANSWER: Put your answer here. [Word count: XX]* 


## Q9

```{r q9codeprovided, warning=FALSE, message=FALSE}
# get the lowest score
lowest <- min(d$audience,na.rm=TRUE)
lowest

# get the highest score
highest <- max(d$audience,na.rm=TRUE)
highest
```

**Q9a**

```{r q9a, warning=FALSE, message=FALSE}
# Put your code here

```

*ANSWER: The probability of the lowest score is ____ and the probability of the highest score is ____.*


**Q9b** 

```{r q9b, warning=FALSE, message=FALSE}
# this code is given
df <- df %>%
  mutate(prob = pnorm(duration,mean=6.5,sd=3))

# you may add additional code here if it's useful to answer the question

```

*ANSWER: Put your answer here. [Word count: XX]* 


**Q9c**

*ANSWER: Put your answer here. [Word count: XX]* 


## Q10

**Q10a**

*ANSWER: Put your answer here. [Word count: XX]* 


**Q10b**

*ANSWER: Put your answer here. [Word count: XX]* 


## Q11

*ANSWER: Put your answer here. Does not contribute to your word count limit.*

